---
title: "Amass_tnc_nitro comment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Code


```{r model, echo=TRUE, message=FALSE}
source("functions and packages/startscripts.R")
library(lme4)
library(lmerTest)
library(effects)


photo_chem <- read.csv("calculated data/Amax_chem.csv")
#run volume format func
photo_chem<- vollab_func(photo_chem)
#get better units for plotting
photo_chem$starch_perc <- photo_chem$starch*100
photo_chem$nitro_mg <- photo_chem$Nmass_notnc *1000


#model and stats
Afit_full <- lmer(A_mass ~ nitro_mg+starch_perc+nitro_mg:starch_perc + (1|ID), data=photo_chem)

# drop interaction 
Afit_noint <- lmer(A_mass ~ nitro_mg + starch_perc + (1|ID), data=photo_chem)

# only N
Afit_N <- lmer(A_mass ~ nitro_mg + (1|ID), data=photo_chem)

# only TNC
Afit_TNC <- lmer(A_mass ~ starch_perc + (1|ID), data=photo_chem)

# Calculate R2
library(MuMIn)
r.squaredGLMM(Afit_full)
r.squaredGLMM(Afit_noint)
r.squaredGLMM(Afit_N)
r.squaredGLMM(Afit_TNC)
```

### **Referee #1**

\
- *Reviewer: 1. The role of nutrient defiency may have been underestimated...The relative contribution should be determined...*

\
*Response*:

We were successfully able to identify not one, but two processes that likely contributed to photosynthetic inhibition. We agree that determining the contribution of both of the processes, nutrient limitation and starch accumulation, is important as they are addressed at length in the discussion. The full model including both nitrogen and starch explains 36% of the variation in *A*~max~ on a mass basis. Including only nitrogen in the model explains 25% and with only starch explains 11% of the variation. The interaction between nitrogen and starch explained practically no variation. Consequently, nutrient deficiency composes 69% of the explained variation and starch accumulation the rest. We have added these new results in the "Gas exchange and photosynthetic parameters" section of the Results. By providing convincing empirical evidence of these processes, this experimental design should provide a framework for future work to further quantify the contributions of each.

\
*Results section*:

The relationship between leaf N~f~ and leaf starch with *A*~max~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). Leaf N~f~ and leaf starch explained 25% and 11% of the variation in *A*~max~ on a mass basis, respectively.


