---
title: "Effects of below-ground space limitation on performance of Eucalyptus seedlings:  Does photosynthesis really control growth?"
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: tree-physiology.csl
bibliography: references.bib
---

\
Courtney E. Campany^1^, Belinda Medlyn^1^, Remko A. Duursma^1^.

\
^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia

\
*Corresponding author*: 
Courtney Campany
E: courtneycampany@gmail.com


```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

```{r global data/sourcing, echo=FALSE, message=FALSE, warning=FALSE}
# Two useful packages for printing numbers, accessing content of objects.
library(broom)
library(reporttools)
# Set Flag to include figures in text or not.
includeFigs <- TRUE

source("functions and packages/startscripts.R")
source("functions and packages/gamplotfunctions.R")

```

```{r, echo=FALSE}
airvarsfig <- "1"
allomfig <- "2"
allocfig <- "3"
asatfig <- "4"
achemfig <- "5"
massmodelfig <- "6"
massmodelfig2 <- "S1"
```


# Abstract {.unnumbered}

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity in order to assess C allocation and biomass partitioning. This study used manipulations of soil volume to test how growth is coupled to physiology, allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown seedlings (free). Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that soil volume effect would be largest in the smallest containers, resulting in physical constraints to growth independently of photosynthesis (A). Photosynthesis would then become sink-limited, resulting in the build-up of leaf nonstructural carbohydrates eventually leading to photosynthetic down regulation. We observed a negative container effect on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes mass, partitioning to leaves, stems, roots was conserved after 120 days. Photosynthetic capacity was also significantly reduced in containers, and was related to both leaf nitrogen content and starch accumulation. We developed a seedling growth model that utilized leaf A rates to allocate daily C uptake towards mass growth of stems, leaves and roots. We then asked whether the observed reductions in A explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in the down regulation of A, these reductions were not significant enough to explain observed growth responses. Thus, as photosynthesis and growth were not coordinated an excess pool of non-biomass C resulted in seedlings with soil volume restriction. This research highlights the need to further utilize mass balance approaches when evaluating plant C allocation and confirms that A and growth are not always directly related.

# Keywords {.unnumbered}
photosynthesis, growth, sink regulation, carbon allocation, soil volume

# Introduction

##Growth and photosynthesis are not always correlated when either sinks (controlling G) or photosynthesis is manipulated.

Should be more about allocation. You mention this but carbon balance must always be respected, so it is about where does the carbon end up (and what do we normally assume about this), not so much about whether A 'controls' G or vice versa. This is certainly how much of the literature has been phrased but you should comment on that, not follow that. 

##This begs the question where the carbon ends up; clearly we must still have carbon balance

##Let's do another manipulative experiment where we affect the sink because we know little about carbon allocation responses to source or sink manipulations in trees.

##We hypothesized that trees would grow slower in small pots but that growth and photosynthesis would be coordinated mediated by a down-regulation of photosynthesis
This study utilizes a novel field design to investigate the coordination between growth and A in *Eucalyptus tereticornis* Sm.~seedlings, by manipulating container size and thus rooting volume. Seedlings were maintained under well watered conditions in order to evaluate only the effect of restricted soil volume and the limited nutrient resource pool. We used freely-rooted seedlings as a control for the container size treatments.  Empirical results were combined with a simple plant growth model to simulate seedling growth with a C mass balance approach, which was then compared to observed harvested seedling mass.  The model used whole-plant C gain, scaled from instantaneous rates of leaf A, to quantify seedling production over 120 days. 

1). The manipulations of container size were expected to induce a belowground sink limitation compared to free seedlings. We hypothesized that declines in seedling grow would be largest in the smallest containers, mediated by a down-regulation of photosynthesis.

2). Resuced sink strength was expected to lead to accumulation of leaf non-structural carbohydrates, which can inhibit photosynthsis. We therefore expected a correlation between carbohydrate accumulation and photosynthetic capacity as a function of soil volume.

3). Last, the growth model was expected to find agreement between simulated and observed seedling mass, through direct correlation of the effects of limiiting soil volume on A.


# Materials and Methods

## Experimental design
This experiment was located at the Hawkesbury Forest Experiment site in Richmond, NSW, Australia. Plots were located in open cover with a site history that consists of a paddock that was converted from native pasture grasses. Top soils at this site, used for the study, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7%) and low water holding capacity. At this site a soil hard layer exists at ~1.0 m with a transition to heavy clay soils. The climate for the region is classified as sub-humid temperate. 

\
*Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain cohort. Previous experiments have confirmed that species with tap roots (similar to *E. tereticornis*) use the center of the container as the medium for thick roots leaving the periphery of the soil as the most active sites for fine root proliferation [@biran1980a; @biran1980b]. This is generally hypothesized to be a different response than seedlings with no taproot. By using a species with tap root growth and manipulations of container length rather than width, it is believed that a more realistic test of inhibition of growth through constrained soil volume would be achieved. Six seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots.

\
Six container volumes were used ranging from 5 l to 35 l, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe and were filled with local top soil (described above). Soil in each container was packed to achieve a target soil bulk density of 1.7 g m^-3^. A Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally sown (free) seedlings. Each experimental block (n=7) contained a complete replicate set of container volumes as well as one free seedling, with 1 m^2^ spacing. For each free seedling, used as the control, a 1 m^2^ subplot was excavated to 0.5 m and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation.

\
Plants were watered weekly or when needed, accounting for natural precipitation, to maintain soil moisture at field capacity (13-15%). Drain systems were built into each pot to prevent pooling of water in containers before root expansion, from reduced root uptake, or from large rainfall events. These conditions could lead to an anaerobic environment around the root that could hinder the uptake of water through reduced root conductance [@poorter2009causes], an undesired experimental artifact. A collection compartment in the bottom of containers, containing gravel covered by root exclusion mesh, was used to collect excess water for 20, 25, and 35 l containers. Plastic tubing (6~mm diameter) was inset into the gravel layer and extended through the top of the container. A lysimeter pump was then used to suction excess water, through the tubing, as needed . As small containers (5, 10, and 15 l) have a larger irradiation effect a simple bottom plug was used to drain excess water from the gravel compartment. 

## Growth and morphology metrics
Seedlings were planted on January 21^st^ 2013 and stem height, diameter at 15~cm and leaf count were measured weekly thereafter. Once the growth rate of individual plants had significantly declined a full biomass harvest was completed (May 21^st^ 2013). Dry mass of leaves, stems, roots and cumulative leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) was measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing cumulative leaf area by total leaf count of only fully expanded leaves. This value was then used to interpolate cumulative leaf area through time with weekly leaf counts. Root mass was collected by passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to 0.5 m depth.  25 g fresh weight subsamples of washed fine roots were analyzed, using WhinoRhizo software (Regent Instruments Inc.), for specific root length (SRL, cm~m^-1^).

## Photosynthetic parameters
Leaf gas exchange measurements were performed bi-weekly at saturating light (A~sat~) and saturating light and [CO~2~] (A~max~) on new fully expanded leaves. Measurements were initiated only after sufficient new leaf growth occurred (March 17^th^, 2013), approximately 6 weeks following planting, and continued until the biomass harvest. Leaf level gas exchange was measured with a standard leaf chamber equipped with blue-red light emitting diodes using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA). A~sat~ measurements were made at PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 400 $\mu$l l^-1^ and A~max~ with [CO~2~] of 1600 $\mu$l l^-1^ and PPFD of 1800 $\mu$mol photons m^-1^ s^-1^. This choice of light level to achieve light saturation is consistent with other studies on *Eucalyptus* species [@kallarackal1997ecophysiological;@pinkard1998photosynthetic;@crous2013photosynthesis;@drake2014capacity]. These measurements were conducted during midday (10:00-14:00 h) with leaf temperature maintained at 25&deg;C. After leaves acclimated to the chamber environment, net CO~2~ assimilation rate and stomatal conductance (g~s~) were logged 5 times for both A~sat~ and A~max~. 

\
Photosynthetic CO~2~ response (AC~i~) curves were also developed at 25&deg;C on a random subset of each container size (n=3) after new leaves were first produced and immediately prior to the final harvest (May 23^rd^ 2013). Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l^-1^ at 25&deg;C at saturating light (above). From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical]. 

\
Leaf dark respiration rates (R~d~) was measured on each seedling during the same dates as AC~i~ curves using detached leaves inside a conifer chamber attached to the Licor 6400 at least 1 hour after sundown.   Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions. Reported values of R~d~ are standardized rates at 25&deg;C using a Q10 value (1.86) developed for these seedlings in a separate experiment (Drake et al. unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.


## Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and R~d~. Leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at midday 13:00-14:30 h. These measurements were used as a measure of static water stress on the seedlings \citep{sellin1999does}, and to ensure that the bulk soil water availability was high enough for plants as they became larger and roots filled the soil volume. 

## Leaf, root and soil chemistry
Leaves used in each gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of N content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer with thermal conductivity and mass spectromic detection (of N~2~ and CO~2~).  The percentage of N in the sample was calculated by comparison with known standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the VPDP scale. Leaf TNC concentration was analyzed using a total starch assay kit (Megazyme International 303 Ireland Ltd., Wicklow, Ireland) and includes the starch (mg g^-1^) and soluble sugar (mg g^-1^) concentrations. Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in [@mitchell2013drought]. Specific leaf area (SLA, m^2^ kg^-1^), for leaves sampled during gas exchange campaigns, was then calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass.

## Seedling growth model

We developed a simple seedling growth model that utilized leaf A rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~i~, stem~i~ and root~i~) and a starting leaf area (LA~i~ measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (P~i~) is then given by

\
(1)
$$P_i = L\left(\frac{C_{day'i}~\sigma_s}{\epsilon_c}\right) - R$$
where L is total plant leaf area (m^2^), C~day,i~ is the predicted daily carbon assimilation (g d^-1^), $\sigma$~s~ is a self shading parameter, $\textepsilon$~c~ is a biomass conversion efficiency parameter and R is the mass based total respiration of all tissue components. Total respiration was calculated as 

\
(2)
$$R = \Sigma(R_c~M_c)$$
where R~c~ is tissue respiration of fine roots, coarse roots or stems (g C g mass^-1^) and M~c~ is the standing biomass of each component. R~leaf~ is represented in the calculation of C~day~ (described below). The change in individual component biomass (M~c~), here solved on a daily time step, is given by

\
(3)
$$\frac{dM_c}{dt} = A_{c}~P_{i}~-~(\Lambda_{c}~M_{c})$$
where A~c~ is the component specific biomass partitioning  to whole plant biomass (%) and $\Lambda$~c~ is component specific turnover rate. Due to the duration of the experiment $\Lambda~stem~ was assumed to equal 0.  Total seedling biomass, per time step, was then equal to the sum of all biomass components; leaves, stems, fine roots and coarse roots. 

\
C~day~ was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] in the 'plantecophys' package in R [@Duursma2014] with the mean photosynthetic parameters (J~max~, V~cmax~, R~d~ and g~1~) for each treatment and meteorological data from an onsite weather station. J~max~ and Vc~max~ were estimated from AC~i~ curves (explained above), R~d~ was emprically measured and the g~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2012reconciling] with observed g~s~ values. Examples of the photosynthesis model are described in Medlyn et al. [-@medlyn2002temperature] and the approach of the coupled leaf gas exchange model are described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15~m intervals, leaf A rates ($\mu$mol~CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. C~day~ was calculated by converting predicted rates to mass C gain over 15~min timesteps (g~m^-2^) and then summed for 24 h. This resulted in 120 unique values of C~day~ for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of C~day~ that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment. 

\
It was further necessary to calculate a self-shading parameter ($\sigma$~s~) when scaling leaf A with total plant leaf area. This was accomplished by utilizing 61 previously digitized Eucalyptus seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package in R (fix cite YplantQMC) to build a 3d plant structure based on digitized metrics of plant allometry and crown structure. Inputting the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total A, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf.  The ratio of total A with self-shading to horizontal leaf was then used to calculate $\sigma$~s~ for each of the 61 digitized seedling, independently for each treatment. Next, the linear relationship between $\sigma$~s~ and total leaf area was for determined across digitized seedlings, within each treatment. For the growth model, $\sigma$~s~ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to C~day,i~. All default parameters used in model simulations are reported in Table.~\ref{table:Table3}.

##Data analysis
Differences in experimental parameters with soil volume were analysed by one-way analysis of variance (ANOVA) in R [@RDevelopment] with individual containers as random effects and soil volume as a categorical fixed effect. Tukey’s post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among soil volume treatments were different. Mixed model ANOVAs of A~max~ and leaf chemistry were performed using the 'nlme' package [@nlme] in R and R^2^ values of mixed models were computed as in [@nakagawa2013general]. Tests of allometric relationships between biomass components were implemented using major axis regression in the 'smatr' package in R [@warton2012smatr]. Results were considered significant at P&leq;0.05..

# Results

## Growth and morphology metrics
In this field study, colder temperatures and reductions in cumulative PPFD per day (Fig. `r airvarsfig`) most likely lead to the reduced growth in the free seedlings in the final weeks of the experiment (Fig. `r allomfig`).  Combined with severe growth reductions in the smallest container volumes the experiment was chosen to be harvested after 120 days. Over this duration height, diameter, and leaf area diverged between container volumes (Fig. `r allomfig`).  First, seedling leaf area significantly diverged between soil volumes (P<0.026) during the 5^th^ week of the experiment. Following this period both height (8^th^} week) and then diameter (9^th^ week) significantly deviated across soil volumes (P<0.002 & 0.001, respectively).  Negative growth effects then manifested as severely reduced height gain and declining leaf area through time with small soil volumes across the final two months of the experiment. Seedlings maintained diameter growth throughout the experiment, although marginal with smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (P<0.001).  Increases in both final stem diameter (P<0.001) and cumulative leaf area (both P<0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling.

\
Total seedling biomass at harvest was significantly different across container volumes (P<0.001) and with free seedlings (P<0.001, Table. 1). We analyzed the relationship between biomass growth with each fold increase in soil volume and found an increase of 34% with a doubling of pot size, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. Additionally, plant biomass was highly correlated with total leaf area across all treatments (R^2^ = 0.97, P<0.001). Differences in biomass partitioning to leaves, stems, and roots were not different across soil volumes after variation in seedling biomass within treatments was factored in the analysis (Fig.`r allocfig`a,b). Across all treatments, the final harvested root:shoot was conserved in these seedlings, with a slightly higher shoot than root mass on average ($\overline{x}$=0.904, Fig. `r allomfig`c).

\
SRL of harvested fine roots was not different across soil volumes (Table. 1). Over the duration of the experiment SLA was higher in free seedlings but was not different across containers sizes (Table. 1, P<0.001) and this pattern was evident in the first gas exchange measurement campaign (P<0.001).

## Leaf chemistry
Leaf N % was significantly higher in free seedlings and the largest container volume at the onset of gas exchange measurements (6th week, P<0.001).  Over the remaining duration of the experiment the smallest container volume had a significant reduction in leaf N % compared to other soil volumes, while free seedlings maintained the highest leaf N % (Table. 1, P<0.001).  Leaf starch content in the smallest container was double that of free seedlings (P=0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table. 1).  Differences in leaf starch between the free seedling and the smallest container were also evident during the first gas exchange campaign (P=0.0013). 

## Gas exchange and photosynthetic parameters
A~sat~ and A~max~ were both significantly higher in the largest container volume and the free seedling at the first measurement campaign (both P<0.001). Across all measurement campaigns A~sat~ (Fig. `r asatfig`) and A~max~ (Table 2.) were consistently higher in free seedlings than in containers (boht P<0.001). The interaction between photosynthetic capacity, leaf starch, and leaf N on a mass basis was marginally significant (P=0.0584) but A~max~ was highly correlated to both leaf N content and leaf starch (both P<0.001). Across all measurement campaign A~max~ was higher when foliar N was also higher, usually associated with low levels of leaf starch (Fig. `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Fig. `r achemfig`b)

\
The photosynthetic parameters J~max~ and Vc~max~ were not different within treatments at the beginning and end of gas exchange campaigns, therefore the parameter means per treatment are reported here (Table. 2).  Overall, both J~max~ and Vc~max~ were significantly higher in free seedlings with little variation between soil volume treatments (P=0.0012 & 0.0021, respectively). Leaf dark respiration rates were not significantly different across soil volumes (Table. 2). The g~1~ parameter, generated for each seedling from the Medlyn et al [-@medlyn2012reconciling] optimal stomatal conductance model, was lowest in the free seedling and was marginally different across soil volume treatments (Table. 2). Predicted values of g~s~, using the g~1~parameter, where highly correlated with observed values (R^2^= .74, P<0.001, data not shown).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 mPa across all seedlings, respectively. Although P=0.001), the mean rates for all seedlings were high at 0.37 mol H~2~0 m^-2^ s^-1^ and did not decline significantly across the experiment duration. Additionally, leaf {$\delta$^13^C at final harvest was not different across treatments (Table. 1). Combined these indices provide strong evidence that water stress was not apparent on these well-watered seedlings throughout the experiment. Soil N % at harvest was not different across soil volumes ($\bar{x}$=04.5 %) and decreased approximately 3 % across all containers over the experiment duration. This indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments. 

## Modelling seedling biomass
Model M~0~, was able to converge on an optimum leaf mass fraction (LMF, 21.6 \%) which predicted the mean harvest total biomass of free seedlings within 1.2 %. Using this optimized LMF, the total biomass of modeled seedlings with soil volume restriction (M~1~) was on average 23&pm;2.4 g C more than measured seedlings when comparing against predicted total net leaf C gain (Fig. `r massmodelfig`a). Thus, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the reductions in daily C assimilation per treatment (Fig. `r massmodelfig`b). As a result, the reductions in leaf A were not sufficient enough to explain the reduction in harvested seedling biomass with soil volume restriction. This resulted in an unexplained pool of excess C, generated from A, that was not allocated to biomass. 

\
The remaining model simulations tested possible C allocation scenarios to account for this excess pool of C. Constraining the model by treatment-specific LMF from the final harvest (M~2~) still overestimated seedling total C by 32&pm;11.1 %, but provided the most improved model predictions (Fig. `r massmodelfig2`a). Using harvested LMF, however, does not capture the increase in senescence of seedlings in small containers (Fig. `r allomfig`c). Thus, the 7-61 % reduction in measured harvested leaf mass fraction from large to small soil volumes compared to the optimized LMF (M~0~) represents an underestimation of realized leaf C allocation. Increases of 50 % in non-leaf tissue respiration (M~3~) improved biomass estimates slightly but overestimated mass C by an average of 46&pm;9.3 % in seedlings with soil volume restriction (Fig. `r massmodelfig2`b). It was further determined that non-tissue respiration rates in modeled seedlings would need to increase by ca. 250 % to account for this entire pool of C. 


# Discussion
As issues still remain regarding how source-sink coordination affects growth of woody species, this study utilized a simple but novel field design to manipulate belowground sink limitations in order to physically restrict *Eucalyptus tereticornis* seedling production. We then addressed questions regarding the coordination of A and growth by complementing empirical results with modelling approaches. We found that reductions in leaf A rates were not sufficient to explain observed reductions in total plant biomass production. We thus encourage the utilization of mass balance approaches and provide direction for future studies when testing factors that control of plant growth under environmental change.





##model setup
I think the point here is that the model is quite standard and classical. It has growth driven by fluctuations in photoynthesis but respiration and allocation are very fixed processes. This reflects much of the growth&production modelling work out there. It turns out that in your experiment this did not really work, which means we must take a closer look at allocation and respiration (including root exudation)
##growth and photosythesis not coordianted as expeceted, consequences for studies of biomass allocation

# Tables {.unnumbered}

```{r table1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
library(ascii)
pve_data <- read.csv("master_scripts/pve_table1.csv")

colnames(pve_data) <- c("Volume (L)", 
                        "Seedling~mass~(g)", 
                        "SLA\\textsubscript{TNC-free}~(m\\textsuperscript{2}~kg\\textsuperscript{-1})",
                        "Leaf~Starch~(\\%)", 
                        "Leaf~Sugars~(\\%)", 
                        "Leaf~Nitrogen~(\\%)",
                        "Root~Nitrogen~(\\%)",
                        "SRL~(cm~m\\textsuperscript{-1})",
                        "{Leaf~\\textdelta}\\textsuperscript{13}C~(\\text{\\textperthousand})")
print(ascii(pve_data,include.rownames=FALSE),type='pandoc')


```
**Table 1.** Responses of plant and leaf characterisitics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass, SRL, root nitrogen and leaf $\delta$^13C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

```{r table2, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
library(ascii)
phys_data <- read.csv("master_scripts/pve_table2.csv")
colnames(phys_data) <- c("Volume~(L)", 
                        "\\textit{A}\\textsubscript{max}", 
                        "\\textit{R}\\textsubscript{dark}",
                        "\\textit{J}\\textsubscript{max}",
                        "\\textit{Vc}\\textsubscript{max}",
                        "\\textit{g}\\textsubscript{s}",
                        "\\textit{g}\\textsubscript{1}")
print(ascii(phys_data,include.rownames=FALSE),type='pandoc')

```
**Table 2.** Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Units for A~max~ and R~dark~ are $\mu$mol m^-2^ s^-1^ and g~s~ are mol m^-2^ s^-1^, each at at 25&deg;C. Values of A~max~, g~s~ and g~1~ represent overall means across measurement campaigns (n=6). R~dark~, J~max~ and Vc~max~ values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

# Figures {.unnumbered}

```{r airvars, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/airvars_plotting.R")
```  

**Figure 1**. Daily maximum and minimum temperature (a), cumulative daily PPFD (b), and daily maximum vapour pressure deficit (c) across the experiment duration in 2013.

```{r allometry, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/allometryplotting.R")
```  

**Figure 2**. Soil volume treatment means &pm; standard error of height growth (a), diameter growth (b), and interpolated seedling leaf area (c) measured weekly of *Eucalyptus tereticornis* seedlings across the experiment duration in 2013.

```{r allocation, fig.width=6, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE}
source("master_scripts/allocation_plotting.R")
```  

**Figure 3**. Soil volume treatment means of mass partitioning to leaves, stems, and roots at harvest (a), bi-variate relationships between mass allocation to leaves and stems + roots (b) and leaf mass as a function of fine root biomass with &pm; standard error (c). For (b) lines represent standardized major axis fitting of the log transformed allometric relationships of leaf mass fraction by treatment. For (c) the dashed line is the 1:1 relationship and the solid line represents the significant linear model fit (R^2^=0.81).

```{r asat, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE}
source("master_scripts/A_stats_plotting.R")
```  


**Figure 4**. Soil volume treatment means &pm; standard error, across all measurement campaigns (n=6), of light saturated rates of photosynthesis at 25&deg;C. Different letters represent significant differences between treatments.

```{r A_tncnitro, fig.width=8, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE}
source("master_scripts/Amax_TNC_N_plotting2.R")
```  

**Figure 5**. Photosynthetic capacity, on a leaf mass basis, as a function of accumulation of leaf starch (a) and leaf nitrogen content without TNC (b).  Colors represent bins levels (n=5) of both leaf starch and nitrogen grouped from low to high .  Lines represents predictions, for each bin level, from the linear mixed effects model equation of A~max~ as a function of starch and nitrogen. The marginal R^2^ (fixed effects only) was 0.37 and the conditional R^2^ (fixed and random effects) was 0.48 for the complete model.

```{r massmodel, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/massmodel_optimize.R")
```  

**Figure 6**. Total carbon mass for harvested and modeled seedlings versus predicted total carbon gain after 120 days (a) and  reductions in final seedling carbon mass, both modeled and observed, as a function of the reduction in leaf photosynthesis across treatments (b). For (a) the dashed 1:1 identifies the difference between net total leaf carbon gain and gross seedling production. For (b) both seedling carbon mass and daily carbon assimilation were first scaled to the free seedling control.

# Supporting Information {.unnumbered}

```{r model_scenarios, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/massmodel_optimize_scenarios.R")
```  

**Figure S1**. Sensitivity testing of seedling growth model to different carbon allocation strategies including; constraints of leaf mass fraction to treatment specific final harvest values (a) and increases in respiration of non-leaf tissue components by 50% (b).  Open and filled symbols represent default model and harvest values, while shaded symbols represent model sensitivity to each scenario by soil volume treatment. Both seedling carbon mass and daily carbon assimilation were first scaled to the free seedling control.

**Table S1** Seedling Growth Model Default Parameters

# References {.unnumbered}

